---
title: "Nectar Pot Number"
author: "Guy Mercer"
date: "13/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
pot_num <- read.csv("input/nectar_pot_number.csv")

work_num <- read.csv("../worker_number/input/worker_number.csv")

pot_num$worker_number <- work_num$worker_number
```

Remove colony 57 due to early queen death and sort variable classes.

```{r}
pot_num <- pot_num [pot_num$colony_number != "57", ]

pot_num$block <- as.factor(pot_num$block)

pot_num$treatment <- as.factor(pot_num$treatment)

pot_num$triad <- as.factor(pot_num$triad)

pot_num$experiment_census_day <- as.numeric(pot_num$experiment_census_day)

pot_num$wax_moth <- as.factor(pot_num$wax_moth)

pot_num$nectar_pot_number <- as.numeric(pot_num$nectar_pot_number)

pot_num$campus_location <- as.factor(pot_num$campus_location)

pot_num$number_of_workers_at_exposure_start <- as.numeric(pot_num$number_of_workers_at_exposure_start)

pot_num$colony_number <- as.factor(pot_num$colony_number)

pot_num$worker_number <- as.numeric(pot_num$worker_number)
```

Look at nectar pot / worker. This shows the highest ratio occur at the end of the colony cycle. Not very informative.

```{r}
# pot_num$ratio <- pot_num$nectar_pot_number / pot_num$worker_number
```

Visualisations

```{r}
library(lattice)

MyLines <- function(xi, yi, ...){
  I <- order(xi)
  panel.lines(xi[I], yi[I], col = 1)
}

# by treatment
xyplot(nectar_pot_number ~ experiment_census_day | treatment, data = pot_num,
       groups = colony_number, xlab = "Day", ylab = "Nectar Pot Number",
       panel = panel.superpose,
       panel.groups = MyLines)
```

I think this plot shows what an unreliable variable nectar pot number is. My process when counting nectar pot number involved only counting those more than half full. However, they all vary in size considerably, meaning the volume of nectar held in each pot is extremely variable. So much so that one large nectar pot could hold the equivalent volume of nectar as five small nectar pots. This underlines the futility of a too in depth analysis into this variable.

Maybe look at the maximum number of nectar pots for each colony. As this is a single point estimate it may be less choppy than the time series data and may indicate whether the insecticide treatments had an effect on foraging capacity. Remove days 0 and 7 as colonies were still in the lab.

For this to work it should have an offset variable for the number of works present at that time.

```{r}
library(tidyverse)

# remove censuses from day 0 and 7.
pot_num <- pot_num [pot_num$experiment_census_day != 0, ]

pot_num <- pot_num [pot_num$experiment_census_day != 7, ]


# determine the max number of nectar pots
col_num <- unique(pot_num$colony_number)

pot_num_max_pot <- tibble()

for (i in 1:length(col_num)) {
  
  by_col <- pot_num [pot_num$colony_number == col_num [i], ]
  
  by_col$max_pot <- max(by_col$nectar_pot_number)
  
  pot_num_max_pot <- rbind(pot_num_max_pot, by_col)
  
}

# use this df.
# max_pot <- pot_num_max_pot [pot_num_max_pot$experiment_census_day == 0,]

# get the max number of nectar pots with the number of workers on that day
library(tidyverse)

max_pot <- tibble()

for (i in 1:nrow(pot_num_max_pot)) {
  
  if(pot_num_max_pot$nectar_pot_number [i] == pot_num_max_pot$max_pot [i]) {
    
    row_match <- pot_num_max_pot [i, ]
    
    max_pot <- rbind(max_pot, row_match)

  }
  

}

# select the earliest day for each colony for when the max pot occurred. 
# at least this is consistent
max_pot_filtered <- tibble()

for (i in 1:length(col_num)) {
  
  by_col <- max_pot [max_pot$colony_number == col_num [i], ]
  
  by_col <- by_col [order(by_col$experiment_census_day),]
  
  by_col <- by_col [1, ]
  
  max_pot_filtered <- rbind(max_pot_filtered, by_col)
  
}

max_pot <- max_pot_filtered
```

Boxplots

```{r}
boxplot(max_pot~treatment,
        varwidth=TRUE,
        data = max_pot)

boxplot(max_pot~block,
        varwidth=TRUE,
        data = max_pot)

boxplot(max_pot~campus_location,
        varwidth=TRUE,
        data = max_pot)
```

Start off with a simple poisson model.

```{r}
library(glmmTMB)

f1 <- formula(max_pot ~ treatment + offset(log(worker_number)) + (1|block) + (1|campus_location))

# poisson
m_poisson <- glmmTMB(f1,
              data = max_pot,
              family = "poisson")

summary(m_poisson)
```

Check for overdispersion using the pearson dispersion statistic.

```{r}
dispfun <- function(m) {
    r <- residuals(m,type="pearson")
    n <- df.residual(m)
    dsq <- sum(r^2)
    c(dsq=dsq,n=n,disp=dsq/n)
}

sapply(list(poisson=m_poisson),dispfun)

```

Model is slightly overdispersed but not by much. How to correct standard errors and p values for this overdispersion.

[How to correct standard errors and p values](https://stackoverflow.com/questions/68915173/how-do-i-fit-a-quasi-poisson-model-with-lme4-or-glmmtmb)

"quasi-likelihood models really represent a post-fitting adjustment to the standard errors of the parameters and the associated statistics; they don't (or shouldn't ...) change anything about the way the model is fitted."

"This way of doing it puts as much of the variance as can be explained by the random effects there, then does a post hoc adjustment for any remaining over (or under)dispersion."

```{r}
quasi_table <- function(model,ctab=coef(summary(model))$cond) {
    phi <- sum(residuals(model, type="pearson")^2)/df.residual(model)
    qctab <- within(as.data.frame(ctab),
    {   `Std. Error` <- `Std. Error`*sqrt(phi)
        `z value` <- Estimate/`Std. Error`
        `Pr(>|z|)` <- 2*pnorm(abs(`z value`), lower.tail=FALSE)
    })
    return(qctab)
}

printCoefmat(quasi_table(m_poisson),digits=4)
```

Just say that you did a Poisson GLM, detected overdispersion, and corrected the standard errors using a quasi-GLM model where the variance is given by φ × μ, where μ is the mean and φ the dispersion parameter.

glmmTMB will fit two parameterizations of the negative binomial: family="nbinom2" gives the classic parameterization with σ2=μ(1+μ/k) (“NB2” in Hardin and Hilbe’s terminology) while family="nbinom1" gives a parameterization with σ2=ϕμ, ϕ>1 (“NB1” to Hardin and Hilbe). The latter might also be called a “quasi-Poisson” parameterization because it matches the mean-variance relationship assumed by quasi-Poisson models, i.e. the variance is strictly proportional to the mean (although the proportionality constant must be >1, a limitation that does not apply to quasi-likelihood approaches).

```{r}
# nb1
m_nb1 <- glmmTMB(f1,
              data = max_pot,
              family = "nbinom1")

summary(m_nb1)
```

Residual plots. Nothing jumps out apart from one large outlier, the colony with 36 nectar pots when there were only 27 workers. However, there is no reason for me to remove this observation. Also, removing it doesn't change the conclusions. 

```{r}
fitted <- predict(m_nb1, type = "response")

E <- resid(m_nb1, type = "pearson")

# resid vs fitted
plot(x = fitted, y = E)
abline(0,0)

# resid vs treatment
boxplot(E~treatment,
        varwidth=TRUE,
        data = max_pot)

# resid vs block
boxplot(E~block,
        varwidth=TRUE,
        data = max_pot)

# resid vs campus_location
boxplot(E~campus_location,
        varwidth=TRUE,
        data = max_pot)
```

This method yields a result where the conclusion does not differ from the manually corrected standard errors and z values. Will be easier to extract predicted values from this method so go forward with it.

Predictions for the offset variable model

```{r}
# create a new df
newdata_df <- data.frame(treatment = c("control", "flup", "sivanto"),
                         block = rep(1,3),
                         campus_location = rep("M", 3),
                         worker_number = c(mean(max_pot$worker_number [max_pot$treatment == "control"]),
                                           mean(max_pot$worker_number [max_pot$treatment == "flup"]),
                                           mean(max_pot$worker_number [max_pot$treatment == "sivanto"])))

# always extract predictions on the link scale process accordingly. 
predictions <- data.frame(predict(m_nb1, newdata_df, type = "link", re.form = NA, se.fit = TRUE))

ilink <- family(m_nb1)$linkinv

predictions <- mutate(predictions,
                fit_resp  = ilink(fit),
                right_upr = ilink(fit + (1.96 * se.fit)),
                right_lwr = ilink(fit - (1.96 * se.fit)))

pred_by_treat <- cbind(newdata_df, predictions)
```

```{r}
# model output plot

model_output <- ggplot() +
  
  geom_point(data = max_pot,
             aes(x = treatment, y = max_pot, color = treatment),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  geom_errorbar(data = pred_by_treat,
                aes(x = treatment,
                    ymin = right_lwr,
                    ymax = right_upr,
                    colour = as.factor(treatment)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = pred_by_treat,
             aes(x = treatment,
                 y = fit_resp,
                 colour = as.factor(treatment)),
             size = 3) +
  
  scale_color_manual(values=c("#117733", "#332288", "#AA4499"),
                     name="Treatment",
                     labels = c("Control", "FPF", "Sivanto")) +
  
  theme_bw() + 
  
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  
  labs(y = "Maximum Nectar Pot Number", x = "Treatment") +
  
  scale_x_discrete(labels = c("Control", "FPF", "Sivanto")) +
  
  geom_segment(aes(x=1.1, xend=2.9, y=35, yend=35)) +
  
  geom_segment(aes(x=1.1, xend=1.9, y=31, yend=31)) +
  
  annotate("text",
           x = c(1.5, 2),
           y = c(32, 36),
           label = c("p = 0.864","p = 0.347"),
           family = "", fontface = 3, size=4) +
  
  ggtitle("Maximum nector pot number for each treatment group when the offset variable worker number is set to its mean value for that treatment group")
  
  # annotate(geom = "text", x = 2.65, y = 37, label = "bold('Control vs Sivanto,')~bolditalic('p')~bold('= 0.0207')", fontface = 2, size = 6, parse = TRUE)

  model_output
```

I think the right thing to do when predicting from this model is to set the offset variable to the mean of that class. The max_nectar_pot predictions are pretty close the mean values. For control (mean = 14.7, pred = 14.04), flup (mean = 14.84, pred = 15.01), sivanto (mean = 10, pred = 10.46). I think the discrepancies are attributable to the predictions ignoring the random effects.

I don't think plots by block or campus location are really necessary. Furthermore, I am unsure how important this metric (max_nectar_pot) really is. 

Before adding the offset variable this was the interpretation:

"For sivanto, colonies displayed a lower maximum capacity for nectar storage, which may indicate lower colony level foraging. Another more conservative interpretation is because nectar pots were a broad range of sizes, and therefore were actually a poor proxy for total stored nectar volume per colony, which is what we are actually interested in, the statistical difference returned probably doesn't represent a true difference in forage storage between treatment groups."

Now the interpretation is simpler. There was no evidence of a difference between maximum nectar pot number between treatment groups once colony size, estimated by worker number, was taken into account.

-----------------------

Below is how to predict from a model with no offset. Slightly different coding so may be useful in future.

```{r, eval=FALSE}
# prediction chunk for model with no offset
# always extract predictions on the link scale process accordingly. 
predictions <- data.frame(predict(m_nb1, type = "link", re.form = NA, se.fit = TRUE))

ilink <- family(m_nb1)$linkinv

predictions <- mutate(predictions,
                fit_resp  = ilink(fit),
                right_upr = ilink(fit + (1.96 * se.fit)),
                right_lwr = ilink(fit - (1.96 * se.fit)))

max_pot <- cbind(max_pot,predictions)

pred_by_treat <- max_pot [2:4, ]

```

```{r, eval=FALSE}
# model output plot for model with no offset
model_output <- ggplot() +
  
  geom_point(data = max_pot,
             aes(x = treatment, y = max_pot, color = treatment),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  geom_errorbar(data = pred_by_treat,
                aes(x = treatment,
                    ymin = right_lwr,
                    ymax = right_upr,
                    colour = as.factor(treatment)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = pred_by_treat,
             aes(x = treatment,
                 y = fit_resp,
                 colour = as.factor(treatment)),
             size = 3) +
  
  scale_color_manual(values=c("#117733", "#332288", "#AA4499"),
                     name="Treatment",
                     labels = c("Control", "FPF", "Sivanto")) +
  
  theme_bw() + 
  
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  
  labs(y = "Maximum Nectar Pot Number", x = "Treatment") +
  
  scale_x_discrete(labels = c("Control", "FPF", "Sivanto")) +
  
  geom_segment(aes(x=1.1, xend=2.9, y=35, yend=35)) +
  
  geom_segment(aes(x=1.1, xend=1.9, y=31, yend=31)) +
  
  annotate("text",
           x = c(1.5, 2),
           y = c(32, 36),
           label = c("p = 0.8067","p = 0.0146"),
           family = "", fontface = 3, size=4)
  
  # annotate(geom = "text", x = 2.65, y = 37, label = "bold('Control vs Sivanto,')~bolditalic('p')~bold('= 0.0207')", fontface = 2, size = 6, parse = TRUE)

model_output
```

```{r}
library(bookdown)

knitr::write_bib(c(.packages(), "bookdown"), "packages.bib")
```

