---
title: "Insecticide Consumption"
author: "Guy Mercer"
date: "13/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Import

```{r}
insecticide_consumption <- read.csv("./input/total_insecticide_dose.csv")
```

Remember only 39 observations as the controls have been removed

Ensure variables have correct class. 

```{r}
library(tidyverse)

# change name to make it easier to recycle code from other markdowns
colnames(insecticide_consumption) [colnames(insecticide_consumption) == "total_insecticide_dose"] <- "summed_consumption"

# make sure variable classes are correct
insecticide_consumption$block <- as.factor(insecticide_consumption$block)
class(insecticide_consumption$block)

insecticide_consumption$treatment <- as.factor(insecticide_consumption$treatment)
class(insecticide_consumption$treatment)

insecticide_consumption$triad <- as.factor(insecticide_consumption$triad)
class(insecticide_consumption$triad)

class(insecticide_consumption$summed_consumption)

insecticide_consumption$number_of_workers_at_exposure_start <- as.numeric(insecticide_consumption$number_of_workers_at_exposure_start)
class(insecticide_consumption$number_of_workers_at_exposure_start)
```

Sort out the cleveland dotplot encoding. 

```{r}
# for a cleveland dotplot to work treatment has to be coded 1-2. 
insecticide_consumption$clevelandcode <- 0 

for (i in 1:nrow(insecticide_consumption)) {
  
  if (insecticide_consumption$treatment [i] == "flup") {
    
    insecticide_consumption$clevelandcode [i] <- 1
    
    }
  
  if (insecticide_consumption$treatment [i] == "sivanto") {
    
    insecticide_consumption$clevelandcode [i] <- 2
    
    }
  
}

# should be numeric already anyway
insecticide_consumption$clevelandcode <- as.numeric(insecticide_consumption$clevelandcode)
```

Use a cleveland dotplot to identify any obvious outliers. 

```{r}
dotchart(insecticide_consumption$summed_consumption,
         groups = factor(insecticide_consumption$clevelandcode),
         ylab = "Order of observations",
         xlab = "Consumption (µg)", main = "Cleveland dotplot", pch = insecticide_consumption$clevelandcode)
```

There does not appear to be any obvious outliers. Maybe consumption is slightly lower for sivanto than the other treatment group?

Have a look at some boxplots

```{r}
boxplot(summed_consumption ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of Consumption Conditional on Treatment", 
        ylab = "Consumption", data = insecticide_consumption)
```

Also want to show consumption vs number_of_workers_exposure_start for each treatment.

```{r}
# by treatment
coplot(summed_consumption ~ number_of_workers_at_exposure_start | treatment, data = insecticide_consumption)

plot(x = insecticide_consumption$number_of_workers_at_exposure_start,
     y = insecticide_consumption$summed_consumption,
     xlab = "starting_worker_number",
     ylab = "insecticide_consumption")
```

I'm going to exclude C17 again as the queen died early on in the exposure regime and it looks like an outlier. Fit beyond optimal model and look at residual plots.

```{r}
# remove C17
insecticide_consumption <- insecticide_consumption [insecticide_consumption$colony_number != "17", ]

# fit beyond optimal model with gls
library(nlme)

# use this for the better resid vs fitted plot
M1 <- lm(summed_consumption ~ treatment + number_of_workers_at_exposure_start + treatment:number_of_workers_at_exposure_start,
          data = insecticide_consumption)

# unhash this for the summary
summary(M1)

# extract residuals from this model. At this stage the ordinary residuals are fine.
# once a variance structure has been applied we'll have to use standardised residuals
# where the ordinary residuals are divided by the square root of the variance. 
# NOTE - standardised residuals = normalised residuals = Pearson residuals (if Poisson GLM).
E <- resid(M1)

# plot residual vs fitted
plot(M1, which = c (1))

# plot residual vs treatment
boxplot(E ~ insecticide_consumption$treatment, main = "Treatment")
abline(0, 0)

# plot residual vs number_of_workers_at_exposure_start
plot(x = insecticide_consumption$number_of_workers_at_exposure_start,
     y = E,
     xlab = "starting_worker_number",
     ylab = "Ordinary Residuals")

# plot residual vs number_of_workers_at_exposure_start by treatment
coplot(E ~ number_of_workers_at_exposure_start | treatment,
       ylab = "Ordinary residuals", data = insecticide_consumption)

# redefine model with gls
M1 <- gls(summed_consumption ~ treatment + number_of_workers_at_exposure_start + treatment:number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "REML")

summary(M1)
```

Looks like variance increases with starting worker number. Try varPower and varExp

```{r}
M2 <- gls(summed_consumption ~ treatment + number_of_workers_at_exposure_start + treatment:number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "REML",
          weights = varPower(form =~ number_of_workers_at_exposure_start))

summary(M2)

anova(M1, M2)

M3 <- gls(summed_consumption ~ treatment + number_of_workers_at_exposure_start + treatment:number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "REML",
          weights = varExp(form =~ number_of_workers_at_exposure_start))

summary(M3)

anova(M1, M3)
```

Neither significant but both have slightly lower AICs. look at the residual plots

```{r}
E2 <- resid(M2, type = "normalized")

# plot residual vs fitted
plot(M2, which = c (1))

# plot residual vs treatment
boxplot(E2 ~ insecticide_consumption$treatment, main = "Treatment")
abline(0, 0)

# plot residual vs number_of_workers_at_exposure_start
plot(x = insecticide_consumption$number_of_workers_at_exposure_start,
     y = E2,
     xlab = "starting_worker_number",
     ylab = "Normalised Residuals")

# plot residual vs number_of_workers_at_exposure_start by treatment
coplot(E2 ~ number_of_workers_at_exposure_start | treatment,
       ylab = "Normalised Residuals", data = insecticide_consumption)
```

```{r}
E3 <- resid(M3, type = "normalized")

# plot residual vs fitted
plot(M3, which = c (1))

# plot residual vs treatment
boxplot(E3 ~ insecticide_consumption$treatment, main = "Treatment")
abline(0, 0)

# plot residual vs number_of_workers_at_exposure_start
plot(x = insecticide_consumption$number_of_workers_at_exposure_start,
     y = E3,
     xlab = "starting_worker_number",
     ylab = "Normalised Residuals")

# plot residual vs number_of_workers_at_exposure_start by treatment
coplot(E3 ~ number_of_workers_at_exposure_start | treatment,
       ylab = "Normalised Residuals", data = insecticide_consumption)
```

The cone shape pattern in the residual vs fitted has gone. Pick varPower as the AIC is slightly lower. Check to see if block does anything. 

```{r}
# standardised residuals for M1 vs triad
E_standard <- resid(M1, type = "normalized")

boxplot(E_standard ~ block, data = insecticide_consumption, axes = TRUE,
        cex.axis=0.75,
        ylab = 'Standardized residuals')
abline(0,0)
```

Perform the test to see if the random intercept should be included. 

```{r}
M4 <- lme(summed_consumption ~ treatment + number_of_workers_at_exposure_start + treatment:number_of_workers_at_exposure_start,
              data = insecticide_consumption,
              random =~ 1 | block, method = "REML",
              weights = varPower(form =~ number_of_workers_at_exposure_start))

summary(M4)

anova(M2, M4)
```

No. So M2 is our final random structure. Now let's begin with finding the optimal fixed structure.

```{r}
M2 <- gls(summed_consumption ~ treatment + number_of_workers_at_exposure_start + treatment:number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "ML",
          weights = varPower(form =~ number_of_workers_at_exposure_start))

M2a <- gls(summed_consumption ~ treatment + number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "ML",
          weights = varPower(form =~ number_of_workers_at_exposure_start))

anova(M2, M2a)

# the interaction term is not significant

M2aa <- gls(summed_consumption ~ number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "ML",
          weights = varPower(form =~ number_of_workers_at_exposure_start))

anova(M2a, M2aa)

M2ab <- gls(summed_consumption ~ treatment,
          data = insecticide_consumption, method = "ML",
          weights = varPower(form =~ number_of_workers_at_exposure_start))

anova(M2a, M2ab)
```

```{r}
# final model is M2a. Redefine with REML
M2a <- gls(summed_consumption ~ treatment + number_of_workers_at_exposure_start,
          data = insecticide_consumption, method = "REML",
          weights = varPower(form =~ number_of_workers_at_exposure_start))

summary(M2a)

# with block as random effect
M2a_ran <- lme(summed_consumption ~ treatment + number_of_workers_at_exposure_start,
              data = insecticide_consumption,
              random =~ 1 | block, method = "REML",
              weights = varPower(form =~ number_of_workers_at_exposure_start))

summary(M2a_ran)
```

Have a look at the residual plots.

```{r}
E2a <- resid(M2a_ran, type = "normalized")

# plot residual vs fitted
plot(M2a, which = c (1))

# plot residual vs treatment
boxplot(E2a ~ insecticide_consumption$treatment, main = "Treatment")
abline(0, 0)

# plot residual vs number_of_workers_at_exposure_start
plot(x = insecticide_consumption$number_of_workers_at_exposure_start,
     y = E2a,
     xlab = "starting_worker_number",
     ylab = "Normalised Residuals")

# plot residual vs number_of_workers_at_exposure_start by treatment
coplot(E2a ~ number_of_workers_at_exposure_start | treatment,
       ylab = "Normalised Residuals", data = insecticide_consumption)

# qqplot
qqnorm(M2a)
```

Use M2a_ran for visualisations as this includes block, which is part of the experimental design.

Namely, if you set level to 0 you get the predictions for which the random effects are set to zero. In the case of linear mixed models, these correspond to population predictions.

Also I don't think the varPower variance structure is required

```{r}
# redefine
final_model <- lme(summed_consumption ~ treatment + number_of_workers_at_exposure_start,
              data = insecticide_consumption,
              random =~ 1 | block, method = "REML",
              weights = varPower(form =~ number_of_workers_at_exposure_start))

# summary(final_model)

# calculate standard error for each point. 
# Design matrix for our observations
xmat <- model.matrix(~ treatment + number_of_workers_at_exposure_start, data=insecticide_consumption)

# Regression coefficients
# betahat <- coef(final_model)

# Predictions
insecticide_consumption$predictions <- predict(final_model, level = 0)
# cbind(head(xmat%*%betahat), head(predictions))

# Sigma^
Sigmahat <- vcov(final_model)

# var/cov(beta0 + beta1*X)
varcovEYhat <- xmat%*%Sigmahat%*%t(xmat)

# Pull off the diagonal elements and take their sqrt to 
# get SEs that quantify uncertainty associated with the line
SEline <- sqrt(diag(varcovEYhat))

# Confidence interval for the mean
insecticide_consumption$upconf <- insecticide_consumption$predictions + 1.96 * SEline
insecticide_consumption$lowconf <- insecticide_consumption$predictions - 1.96 * SEline

treatment_levs <- unique(insecticide_consumption$treatment)

treatment_preds_df <- tibble()

for (i in 1:length(treatment_levs)) {
  
  # order predictions according to treatment
  treatment_preds <- insecticide_consumption [insecticide_consumption$treatment == treatment_levs [i],]
  
  # sort values by number of workers size
  treatment_preds <- treatment_preds [order(treatment_preds$number_of_workers_at_exposure_start),]
  
  treatment_preds_df <- rbind(treatment_preds_df, treatment_preds)
  
}

model_preds <- ggplot() + 
      geom_point(data = insecticide_consumption, aes(x = number_of_workers_at_exposure_start, y = summed_consumption, color = treatment)) +
      geom_line(data = treatment_preds_df, aes(x = number_of_workers_at_exposure_start, y = predictions, group = treatment, color = treatment)) +
      geom_ribbon(data = treatment_preds_df, aes(x = number_of_workers_at_exposure_start, y = predictions, ymin = lowconf, ymax = upconf,
                                              group = treatment, color = treatment, fill = treatment), alpha = 0.3) +
      theme_bw() +
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      scale_color_manual(values=c("#332288", "#AA4499"), name = "Treatment", labels = c("FPF", "Sivanto")) +
      scale_fill_manual(values=c("#332288", "#AA4499"), name = "Treatment", labels = c("FPF", "Sivanto")) +
      xlab("Starting Worker Number") +
      ylab("Insecticide Consumption (µg)") +
      ggtitle("Aggregate insecticide consumption during the exposure window differed between treatment groups")

model_preds

```

