---
title: "Data Transformation and Cleaning"
author: "Guy Mercer"
date: "13/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
queen_rearing <- read.csv("./input/queen_rearing.csv")

# import another csv with triad, batch etc. information
batch_info_csv <- read.csv("./input/gps_coordinates.csv")
```

From this csv I need to calculate the number of days from capture to

a)  egg laying

b)  > = 6 workers

c)  exposure start

Why? If a queen took a long time to lay eggs and/or form a colony, this could be indicative in some way of lower fitness. Naturally predisposed to producing smaller colonies. Worth checking in any case.

```{r}
library(tidyverse)

# df combining function
col_name_df_combiner_function <- function(dataframe_1, dataframe_2) {
        
    combined_df1 <- tibble()
      
    for (i in 1:nrow(dataframe_1)) {
          
      for (j in 1:nrow(dataframe_2)) {
              
        if (dataframe_1 [i, 1] == dataframe_2 [j , 1]) {
                
          rows_combined <- cbind(dataframe_1 [i, ], dataframe_2 [j, ])
                    
          combined_df1 <- rbind(combined_df1, rows_combined)
                
        }
              
      }
            
    }
        
    combined_df1 <- combined_df1 [, unique(colnames(combined_df1))]
    
    return(combined_df1)
        
  }
# explanatory_variables function
explanatory_variables_function <- function(queen_rearing, batch_info_csv) {
  
  # only retain 60 colonies that made it to exposure start. 
  treatments <- c("control", "sivanto", "flup")
  
  treatment_assigned_only <- tibble()
  
  for (i in 1:length(treatments)) {
    
    treatment_group_filter <- queen_rearing [grep(treatments [i], queen_rearing$treatment_assigned), ]
    
    treatment_assigned_only <- rbind(treatment_assigned_only, treatment_group_filter)
    
  }
  
  # sort out date format so it can be interpreted by R.
  # date_caught <- as.Date(treatment_assigned_only$date_caught, format = "%d.%m.%y")
  
  # remove_date_of_queen_death_column
  treatment_assigned_only <- treatment_assigned_only [, -grep("queen_death", colnames(treatment_assigned_only))]
  
  # columns in date format
  date_columns <- grep("date", colnames(treatment_assigned_only))
  
  # other useful columns. Retain and put at front of new dataframe.
  formatted_date_categories <- tibble(treatment_assigned_only [, c(grep("colony_number", colnames(treatment_assigned_only)), 
                                                                   grep("rearing_location", colnames(treatment_assigned_only)),
                                                                   grep("number_of_workers_at_exposure_start", colnames(treatment_assigned_only)),
                                                                   grep("treatment_assigned", colnames(treatment_assigned_only))
                                                                   )])
  
  # extract columns in date format, convert to date and add to new dataframe with appropriate col name. 
  for (i in 1:length(date_columns)) {
    
    indiv_date_cats <- as.Date(treatment_assigned_only [, date_columns [i]], format = "%d.%m.%y")
    
    formatted_date_categories <- cbind(formatted_date_categories, indiv_date_cats)
    
    colnames(formatted_date_categories) [i+4] <- colnames(treatment_assigned_only [date_columns [i]])
  
  }
  
  # remove date columns of either no use in later analysis (date of second screen pass and date of colony box move)
  # or measures that were compromised by time pressure (date of first worker).
  # I could have used the col indexes directly but this way there is no error for mistyping. 
  unwanted_cols <- c("date_of_second_screen", "date_of_first_worker", "date_of_colony_box_move")
  
  for (i in 1:length(unwanted_cols)) {
    
    formatted_date_categories <- formatted_date_categories [, -grep(unwanted_cols [i], colnames(formatted_date_categories))]
    
  }
  
  # work out time_to_egg_laying, time_to_>=_6_workers, time_to_exposure_start
  date_calc <- colnames(formatted_date_categories) [grep("date", colnames(formatted_date_categories))]
  
  new_col_names <- c("time_to_egg_laying", "time_to_6_workers", "time_to_exposure_start")
  
  for (i in 2:length(date_calc)) {
    
    length_of_days <- formatted_date_categories [grep(date_calc [i], colnames(formatted_date_categories))] - formatted_date_categories [grep(date_calc [1], colnames(formatted_date_categories))]
    
    colnames(length_of_days) <- new_col_names [i-1]
      
    formatted_date_categories <- cbind(formatted_date_categories, length_of_days)
  
  }
  
  # change the time_diff class to numeric for newly generated explanatory variables.
  formatted_date_categories$time_to_egg_laying <- as.numeric(formatted_date_categories$time_to_egg_laying)
  formatted_date_categories$time_to_6_workers <- as.numeric(formatted_date_categories$time_to_6_workers)
  formatted_date_categories$time_to_exposure_start <- as.numeric(formatted_date_categories$time_to_exposure_start)
  
  # save date caught separately
  date_caught <- cbind(colony_number = formatted_date_categories$colony_number, date_caught = formatted_date_categories$date_caught)
  
  # remove date columns
  formatted_date_categories <- formatted_date_categories [ , -grep("date", colnames(formatted_date_categories))]
  
  # add other important information to the dataframe like triad, batch etc. 
  
  # removes rows that contain only NAs and blank cells.
  # is.na(batch_info_csv) | batch_info_csv == "" this creates a logical matrix of whether a cell contains NA or is blank
  # !apply(is.na(batch_info_csv) | batch_info_csv == "", 1, all) searches aforementioned logical matrix by row and ascertains where all of the values are true. If they are all true (all NAs or blanks, a FALSE is returned due to the !)
  # batch_info_csv[!apply(is.na(batch_info_csv) | batch_info_csv == "", 1, all),] This logical vector is then applied to the rows of batch_info_csv
  
  batch_info_csv <- batch_info_csv[!apply(is.na(batch_info_csv) | batch_info_csv == "", 1, all),]
  
  # remove blank/NA columns
  batch_info_csv <- batch_info_csv [, !apply(is.na(batch_info_csv), 2, all)]
  
  # remove lat and long
  batch_info_csv <- batch_info_csv [, -grep(pattern = "\\w*tude", x = colnames(batch_info_csv))]
  
  # combine the two dfs using colony number.  
  
  combined_df <- col_name_df_combiner_function(dataframe_1 = batch_info_csv, dataframe_2 = formatted_date_categories)
  
  # make variable naming clearer
  colnames(combined_df) <- sub("^location$", "campus_location", colnames(combined_df))
  
  # create a queen capture date variable. Do not think of it as time. For example, with reproductive output, if two data points are separated by two queen capture days this does not mean that the reproductive output datapoints are separated by two days in time. Therefore, it is not a time series. Treat as a regular numerical, continuous explanatory variable
  queen_capture_day <- cbind(date_caught [, grep("colony_number", colnames(date_caught))], rep(0, 60))
  
  for (i in 1:nrow(date_caught)) {
    
     capture_day <- date_caught [i, grep("date_caught", colnames(date_caught))] - min(date_caught [, grep("date_caught", colnames(date_caught))]) + 1
     
     queen_capture_day [i, 2] <- as.numeric(capture_day)
  
  }
  
  colnames(queen_capture_day) <- c("colony_number", "queen_capture_day")
  
  queen_capture_day <- as.data.frame(queen_capture_day)
  
  final_output <- col_name_df_combiner_function(dataframe_1 = combined_df, dataframe_2 = queen_capture_day)
  
  return(final_output)

}

explanatory_variables <- explanatory_variables_function(queen_rearing = queen_rearing, batch_info_csv = batch_info_csv)
```

The next task is to create the col_survival_days and queen_survival_days variables from the weekly_census_data sheet.

```{r}
# import
weekly_census_data <- read.csv("./input/weekly_census_data.csv")

# remove rows only containing blank cells and n/a
weekly_census_data <- weekly_census_data[!apply(is.na(weekly_census_data) | weekly_census_data == "", 1, all),]

# remove blank/NA columns
weekly_census_data <- weekly_census_data [, !apply(is.na(weekly_census_data), 2, all)]

survival_variable_function <- function(survival_column_name) {
  
  # keep only weeks where colonies were dead
  # eval(parse(paste0())) allows the calling of variables named after the string defined by paste0
  colony_n <- weekly_census_data [eval(parse(text = paste0("weekly_census_data$", survival_column_name, "_alive", sep=""))) == "N", ]
  
  # using unique colony numbers, filter rows matching to said col number, find the smallest experimental_day value for these rows and save both
  # col_number and experimental_day.
  survival_days_df <- tibble()
  
    for (i in 1:length(unique(colony_n$colony_number))) {
      
      individual_colony <- colony_n [unique(colony_n$colony_number) [i] == colony_n$colony_number, ] 
    
      survival <- min(individual_colony$experiment_census_day)
    
      survival_days_df [i, 1] <- unique(colony_n$colony_number) [i]
      
      survival_days_df [i, 2] <- survival
    
    }
  
  colnames(survival_days_df) <- c("colony_number", paste0(survival_column_name,"_survival_days", sep=""))
  
  combined_df1 <- col_name_df_combiner_function(dataframe_1 = explanatory_variables, dataframe_2 = survival_days_df)
  
  combined_df1 <- combined_df1 [, -grep("treatment_assigned", colnames(combined_df1))]

  return(combined_df1)
  
}

queen_survival_days_df <- survival_variable_function(survival_column_name = "queen")

colony_survival_days_df <- survival_variable_function(survival_column_name = "colony")
```

Next create the worker_number and nectar_pot_number sheets with all the relevant variables

```{r}
worker_number_nectar_pot_function <- function(variable_name) {
  
  # keep only the required columns from weekly_census_data
  relevant_variables_weeky_census_data <- c("colony_number", "block", "treatment", "triad", "experiment_census_day", "wax_moth",
                                            paste0(variable_name, "_number", sep=""))
  
  filtered_weekly_census_data <- weekly_census_data [, relevant_variables_weeky_census_data]
  
  # do the same for explanatory_variables
  relevant_variables_explanatory_variables <- c("colony_number", "campus_location", "rearing_location", "number_of_workers_at_exposure_start",
                                                "time_to_egg_laying", "time_to_6_workers", "time_to_exposure_start", "queen_capture_day")
  
  filtered_explanatory_variables <- explanatory_variables [, relevant_variables_explanatory_variables]
  
  # combine the two
  combined_df1 <- col_name_df_combiner_function(dataframe_1 = filtered_weekly_census_data, dataframe_2 = filtered_explanatory_variables)
  
  return(combined_df1)
  
}

worker_number_df <- worker_number_nectar_pot_function(variable_name = "worker")

nectar_pot_number_df <- worker_number_nectar_pot_function(variable_name = "nectar_pot")
```

Next create a male number spreadsheet, followed by one that incorporates gynes by applying a weighting of 2 to them.

```{r}

total_male_number_function <- function(weekly_census_data) {
  
  total_male_num_df <- tibble()

  for (i in 1:length(unique(weekly_census_data$colony_number))) {
    
    by_colony <- weekly_census_data [weekly_census_data$colony_number == unique(weekly_census_data$colony_number) [i], ]
    
    total_male_num <- sum(as.numeric(as.character(by_colony$male_number)))
    
    total_male_num_df [i, 1] <- unique(weekly_census_data$colony_number) [i]
    
    total_male_num_df [i, 2] <- total_male_num
    
  }
  
  colnames(total_male_num_df) <- c("colony_number", "total_male_number")
  
  # combine the explanatory variables
  combined_df1 <- col_name_df_combiner_function(dataframe_1 = total_male_num_df, dataframe_2 = explanatory_variables)
  
  combined_df1 <- combined_df1 [, -grep("treatment_assigned", colnames(combined_df1))]
  
  return(combined_df1)
  
}

total_male_number <- total_male_number_function(weekly_census_data = weekly_census_data)
```

Add a total_reproductive_output (male_number + (gyne_number \* 2)) column to the total_male_number spreadsheet.

```{r}
reproductive_output_function <- function(weekly_census_data, total_male_number) {
  
  gyne_corr_num_df <- tibble()

  for (i in 1:length(unique(weekly_census_data$colony_number))) {
      
    by_colony <- weekly_census_data [weekly_census_data$colony_number == unique(weekly_census_data$colony_number) [i], ]
      
    total_gyne_num <- sum(as.numeric(as.character(by_colony$gyne_number)))
    
    total_male_num <- sum(as.numeric(as.character(by_colony$male_number)))
      
    gyne_corr_num <- total_gyne_num * 2
    
    total_reproductive_output <- gyne_corr_num + total_male_num
      
    gyne_corr_num_df [i, 1] <- unique(weekly_census_data$colony_number) [i]
      
    gyne_corr_num_df [i, 2] <- total_reproductive_output
      
  }
    
  colnames(gyne_corr_num_df) <- c("colony_number", "total_reproductive_output")
    
  # add this column to total_male_number.
  
  reproductive_output_df <- col_name_df_combiner_function(dataframe_1 = gyne_corr_num_df, dataframe_2 = total_male_number)
  
  return(reproductive_output_df)
  
}

total_reproductive_output_df <- reproductive_output_function(weekly_census_data = weekly_census_data, total_male_number = total_male_number)
```

Add a simple binary wax moth variable to total_reproductive_number, colony_survival and queen_survival.

```{r}
wax_moth_variable_function <- function(weekly_census_data, survival_df) {
  
  wax_moth_cols <- weekly_census_data [weekly_census_data$wax_moth == "Y", ]

  wax_moth_cols <- unique(wax_moth_cols$colony_number)
  
  survival_df$wax_moth <- "N"
  
  for (i in 1:length(wax_moth_cols)) {
    
    for (j in 1:nrow(survival_df)) {
      
      if (wax_moth_cols [i] == survival_df$colony_number [j]) {
        
        survival_df$wax_moth [j] <- "Y" 
        
      }
      
    }
    
  }
  
  return(survival_df)

}

colony_survival_days_df <- wax_moth_variable_function(weekly_census_data = weekly_census_data, survival_df = colony_survival_days_df)

queen_survival_days_df <- wax_moth_variable_function(weekly_census_data = weekly_census_data, survival_df = queen_survival_days_df)

total_reproductive_output_df <- wax_moth_variable_function(weekly_census_data = weekly_census_data, 
                                                           survival_df = total_reproductive_output_df)

```

Add queen survival time to total_reproductive_output

```{r}
total_reproductive_output_df <- col_name_df_combiner_function(dataframe_1 = total_reproductive_output_df, dataframe_2 = queen_survival_days_df)
```

Tackle the forage/insecticide consumption response variables. As I have only have 6 time points sum the consumption.

```{r}
# import the csv
forage_consumption <- read.csv("./input/forage_consumption_exposure_period.csv")

starting_worker_number <- explanatory_variables [, grep("colony_number|workers_at_exposure", colnames(explanatory_variables))]

pollen_consumption_function <- function(forage_consumption) {
  
  # remove blank rows
  forage_consumption <- forage_consumption[!apply(is.na(forage_consumption) | forage_consumption == "", 1, all),]
  
  days <- c("0", "1", "2", "4", "6", "8", "10")
  
  pollen_consumption <- tibble(matrix(0 ,nrow = 75, ncol = 1))
  
  # subtract day out columns from day in columns
  for (i in 1:(length(days) -1)) {
    
    pollen_consumption_day <- eval(parse(text = paste0("forage_consumption$pollen_mass_in_day_", days [i], sep=""))) -
                              eval(parse(text = paste0("forage_consumption$pollen_mass_out_day_", days [i+1], sep="")))
  
    pollen_consumption [, i] <- pollen_consumption_day
  
    colnames(pollen_consumption) [i] <- paste0("pollen_consumption_day_", days [i+1], sep="")
    
  }
  
  # associate with colony_number
  pollen_consumption <- cbind(colony_number = forage_consumption$colony_number, block = forage_consumption$block, treatment = forage_consumption$treatment, triad = forage_consumption$triad, pollen_consumption)
  
  # check to see if it has worked properly. Are all the values of the evap controls small?
  ECs <- pollen_consumption [grep("EC", pollen_consumption$colony_number), ]
  
  # check by manually comparing to a few random values.if TRUE it has worked
  (forage_consumption$pollen_mass_in_day_0 [42] - forage_consumption$pollen_mass_out_day_1 [42]) == pollen_consumption [pollen_consumption$colony_number == forage_consumption$colony_number [42], colnames(pollen_consumption) == "pollen_consumption_day_1"]
  
  (forage_consumption$pollen_mass_in_day_0 [56] - forage_consumption$pollen_mass_out_day_1 [56]) == pollen_consumption [pollen_consumption$colony_number == forage_consumption$colony_number [56], colnames(pollen_consumption) == "pollen_consumption_day_1"]
  
  # now correct the values by the mean of the evaporation controls
  corrected_consumption_df <- tibble()
  
  for (i in 1:length(unique(pollen_consumption$block))) {
    # subset by block
    by_block <-  pollen_consumption [pollen_consumption$block == unique(pollen_consumption$block) [i], ]
    
    # keep the evap controls
    evap_controls <- by_block [grep("EC", by_block$colony_number), ]
   
    # keep only the consumption data
    consumption_only <- evap_controls [, grep("consumption", colnames(evap_controls))]
    
    # take an average of the EC values
    evap_control_means <- colMeans(consumption_only)
    
    # correct each column with the mean EC value
    batch_corrected_consumption_df <- tibble(matrix(0, nrow = 12, ncol = 6))
   
    for (j in 1:length(evap_control_means)) {
      
      no_evap_controls <- by_block [-grep("EC", by_block$colony_number), ]
   
      corrected_consumption <- no_evap_controls [, colnames(no_evap_controls) == names(evap_control_means) [j]] - evap_control_means [j]
      
      batch_corrected_consumption_df [, j] <- corrected_consumption
      
      colnames(batch_corrected_consumption_df) [j] <- names(evap_control_means) [j]
    }
    
    # re-associate with identifiers
    batch_corrected_consumption_df <- cbind(colony_number = no_evap_controls$colony_number, block = no_evap_controls$block,
                                      treatment = no_evap_controls$treatment, triad = no_evap_controls$triad, 
                                      batch_corrected_consumption_df)
    
    corrected_consumption_df <- rbind(corrected_consumption_df, batch_corrected_consumption_df)
    
  }
  
  # check to see if it has worked properly manually.
  evap_correction_checker <- function(col_num) {
    
    block_evap_controls <- pollen_consumption [pollen_consumption$block == pollen_consumption [pollen_consumption$colony_number == col_num, ]$block, ] [grep("EC", pollen_consumption [pollen_consumption$block == pollen_consumption [pollen_consumption$colony_number == col_num, ]$block, ] $colony_number), ]
    
    pollen_consumption [pollen_consumption$colony_number == col_num, ] [, names(colMeans(block_evap_controls [, grep("consumption", colnames(block_evap_controls))])) [1]] - colMeans(block_evap_controls [, grep("consumption", colnames(block_evap_controls))]) [1] ==
      corrected_consumption_df [corrected_consumption_df$colony_number == col_num, colnames(corrected_consumption_df) == names(colMeans(block_evap_controls [, grep("consumption", colnames(block_evap_controls))])) [1]]
    
  }
  
  # will return true or false
  evap_correction_checker(col_num = "176")
  
  # set negative values to 0
  for (i in 1:ncol(corrected_consumption_df [, grep("consumption", colnames(corrected_consumption_df))])) {
    
    consumption_column <- corrected_consumption_df [, grep("consumption", colnames(corrected_consumption_df))] [i]
    
    for (j in 1:nrow(consumption_column)) {
      
      if (consumption_column [j, ] < 0 ) {
        
        consumption_column [j, ] <- 0
        
      }
      
    }
    
    corrected_consumption_df [, colnames(corrected_consumption_df) == names(consumption_column)] <- consumption_column
    
  }
  
  # sum for total pollen consumption over exposure_window
  summed_consumption <- as.data.frame(apply(corrected_consumption_df [,grep("consumption", colnames(corrected_consumption_df))], 1, sum))
  
  names(summed_consumption) <- "summed_consumption"
  
  corrected_consumption_df <- cbind(corrected_consumption_df, summed_consumption)
  
    # add starting_worker_number
  corrected_consumption_df1 <- col_name_df_combiner_function(dataframe_1 = corrected_consumption_df, 
                                                             dataframe_2 = starting_worker_number)
  
  return(corrected_consumption_df1)
  
}

pollen_consumption_df <- pollen_consumption_function(forage_consumption = forage_consumption)
```

Do the same for sucrose. There were 4 spills for sucrose so make sure to exclude these at the end.

```{r}
# import the csv
forage_consumption <- read.csv("./input/forage_consumption_exposure_period.csv")

starting_worker_number <- explanatory_variables [, grep("colony_number|workers_at_exposure", colnames(explanatory_variables))]

sucrose_consumption_function <- function(forage_consumption, starting_worker_number) {
  
  # remove blank rows
  forage_consumption <- forage_consumption[!apply(is.na(forage_consumption) | forage_consumption == "", 1, all),]
  
  days <- c("0", "1", "2", "4", "6", "8", "10")
  
  sucrose_consumption <- tibble(matrix(0 ,nrow = 75, ncol = 1))
  
  # subtract day out columns from day in columns
  for (i in 1:(length(days) -1)) {
    
    sucrose_consumption_day <- eval(parse(text = paste0("forage_consumption$sucrose_mass_in_day_", days [i], sep=""))) +
                              eval(parse(text = paste0("forage_consumption$feeder_mass_day_", days [i], sep=""))) -
                              eval(parse(text = paste0("forage_consumption$sucrose_mass_out_day_", days [i+1], sep="")))
  
    sucrose_consumption [, i] <- sucrose_consumption_day
  
    colnames(sucrose_consumption) [i] <- paste0("sucrose_consumption_day_", days [i+1], sep="")
    
  }
  
  # associate with colony_number
  sucrose_consumption <- cbind(colony_number = forage_consumption$colony_number, block = forage_consumption$block, treatment = forage_consumption$treatment, triad = forage_consumption$triad, sucrose_consumption)
  
  # check to see if it has worked properly. Are all the values of the evap controls small?
  # no as there was a spill in the ECs. 
  ECs <- sucrose_consumption [grep("EC", sucrose_consumption$colony_number), ]
  
  # now correct the values by the mean of the evaporation controls
  corrected_consumption_df <- tibble()
  
  for (i in 1:length(unique(sucrose_consumption$block))) {
    # subset by block
    by_block <-  sucrose_consumption [sucrose_consumption$block == unique(sucrose_consumption$block) [i], ]
    
    # keep the evap controls
    evap_controls <- by_block [grep("EC", by_block$colony_number), ]
   
    # keep only the consumption data
    consumption_only <- evap_controls [, grep("consumption", colnames(evap_controls))]
    
    # take an average of the EC values
    evap_control_means <- colMeans(consumption_only)
    
    # correct each column with the mean EC value
    batch_corrected_consumption_df <- tibble(matrix(0, nrow = 12, ncol = 6))
   
    for (j in 1:length(evap_control_means)) {
      
      no_evap_controls <- by_block [-grep("EC", by_block$colony_number), ]
   
      corrected_consumption <- no_evap_controls [, colnames(no_evap_controls) == names(evap_control_means) [j]] - evap_control_means [j]
      
      batch_corrected_consumption_df [, j] <- corrected_consumption
      
      colnames(batch_corrected_consumption_df) [j] <- names(evap_control_means) [j]
    }
    
    # re-associate with identifiers
    batch_corrected_consumption_df <- cbind(colony_number = no_evap_controls$colony_number, block = no_evap_controls$block,
                                      treatment = no_evap_controls$treatment, triad = no_evap_controls$triad, 
                                      batch_corrected_consumption_df)
    
    corrected_consumption_df <- rbind(corrected_consumption_df, batch_corrected_consumption_df)
    
  }
  
  # for colony_34 day_10 two feeders were used. Therefore, two evaporation control corrections should be performed
  # to save time and lots of coding apply the additional correction manually.
  # mean of EC1-3 for day 10 = 796. Subtract this value from colony_34 day_10.
  corrected_consumption_df [corrected_consumption_df$colony_number == "34",
                          colnames(corrected_consumption_df) == "sucrose_consumption_day_10"] <-
    corrected_consumption_df [corrected_consumption_df$colony_number == "34",
                          colnames(corrected_consumption_df) == "sucrose_consumption_day_10"] - 796
  
  # sum for total sucrose consumption over exposure_window
  summed_consumption <- as.data.frame(apply(corrected_consumption_df [,grep("consumption", 
                                                                              colnames(corrected_consumption_df))], 1, sum))
  
  names(summed_consumption) <- "summed_consumption"
  
  corrected_consumption_df <- cbind(corrected_consumption_df, summed_consumption)
  
  # remove the 4 colonies that had spills. 
  sucrose_mass_out_values <- forage_consumption [, grep("sucrose_mass_out", colnames(forage_consumption))]
  
  sucrose_mass_out_values <- cbind(colony_number = forage_consumption [, grep("colony_number", colnames(forage_consumption))], sucrose_mass_out_values)
  
  zero_or_not <- sucrose_mass_out_values == 0
  
  any_zero_in_row <- apply(zero_or_not, 1, any)
  
  spill_rows <- sucrose_mass_out_values [any_zero_in_row,]
  
  spill_rows_col_num <- spill_rows$colony_number
  
  for (i in 1:length(spill_rows_col_num)) {
    
    corrected_consumption_df <- corrected_consumption_df [spill_rows_col_num [i] != corrected_consumption_df$colony_number, ]
    
  }
  
  # add starting_worker_number
  corrected_consumption_df1 <- col_name_df_combiner_function(dataframe_1 = corrected_consumption_df, 
                                                             dataframe_2 = starting_worker_number)
  
  return(corrected_consumption_df1)
  
}

sucrose_consumption_df <- sucrose_consumption_function(forage_consumption = forage_consumption, starting_worker_number = starting_worker_number)
```

Create insecticide_consumption_df

```{r}
# create a wide to long function
wide_to_long_function <- function(dataframe_1, forage_type) {
  
  # convert sucrose_consumption into long format
  days <- c(1, 2, 4, 6, 8, 10)
  
  consumption_long <- tibble()
  
  for (i in 1:length(days)) {
    
    # retain the identifier and explan var columns plus a day sucrose consumption column 
    consumption_by_day <- dataframe_1 [, c(grep("colony_number", colnames(dataframe_1)),
                                                              grep("block", colnames(dataframe_1)),
                                                              grep("treatment", colnames(dataframe_1)),
                                                              grep("triad", colnames(dataframe_1)),
                                                              grep("number_of_workers_at_exposure_start",
                                                                                                       colnames(dataframe_1)),
                                                              grep(paste0("^",forage_type,"_consumption_day_", days [i], "$", sep = ""),
                                                                 colnames(dataframe_1)))]
    
    # add a column with the day
    consumption_by_day$day <- days [i]
    
    # rename the sucrose_consumption column so it has the same name for rbind to work
    names(consumption_by_day) [names(consumption_by_day) == paste0(forage_type,"_consumption_day_", 
                                                                                   days [i], sep="")] <- paste0(forage_type, "_consumption",
                                                                                                                sep = "")
    
    consumption_long <- rbind(consumption_long, consumption_by_day)
    
  }
  
  return(consumption_long)
  
}

sucrose_consumption_long <- wide_to_long_function(dataframe_1 = sucrose_consumption_df, forage_type = "sucrose")

pollen_consumption_long <- wide_to_long_function(dataframe_1 = pollen_consumption_df, forage_type = "pollen")

# convert from mg to kg
sucrose_consumption_long$sucrose_consumption <- sucrose_consumption_long$sucrose_consumption / 1000000
pollen_consumption_long$pollen_consumption <- pollen_consumption_long$pollen_consumption / 1000000

# create an insecticide_dose_function
insecticide_dose_function <- function(dataframe_1, forage_type) {
  
  days <- c(1, 2, 4, 6, 8, 10)
  # these values are in µg/kg (µg insecticide / kg feed)
  sucrose_insecticide_concs <- c(day_1 = 866.2034, day_2 = 514.0244, day_4 = 243.537,
                                 day_6 = 89.95134, day_8 = 33.22388, day_10 = 12.27137)
  
  pollen_insecticide_concs <- c(day_1 = 19187.14, day_2 = 5033.175, day_4 = 689.9583,
                                day_6 = 689.9583, day_8 = 689.9583, day_10 = 689.9583)
  
  # create "empty" insecticide_consumption_column
  dataframe_1$insecticide_consumption <- 0
  
  
  # match the correct concentration using day and multiply this values (µg/kg) by the mass of forage consumed (kg).
  # yields µg of insecticide consumed. 
  for (i in 1:length(days)) {
    
    for (j in 1:nrow(dataframe_1)) {
      
      if (days [i] == dataframe_1$day [j]) {
        
        dataframe_1$insecticide_consumption [j] <- eval(parse(text = paste0("dataframe_1$", forage_type, "_consumption", sep =""))) [j] * 
                                                                eval(parse(text = paste0(forage_type,"_insecticide_concs",
                                                                                         sep = ""))) [i]
        
      }
      
    }
    
  }
  
  return(dataframe_1)
}

sucrose_consumption_long <- insecticide_dose_function(dataframe_1 = sucrose_consumption_long, forage_type = "sucrose")

pollen_consumption_long <- insecticide_dose_function(dataframe_1 = pollen_consumption_long, forage_type = "pollen")
```

```{r}
# create a function that returns total insecticide dose for each colony
total_insecticide_dose_function <- function(forage_consumption, pollen_consumption_long, sucrose_consumption_long, sucrose_consumption_df) {
  
  # remove pollen rows where there was a spill for sucrose
  forage_consumption <- forage_consumption[!apply(is.na(forage_consumption) | forage_consumption == "", 1, all),]

  sucrose_mass_out_values <- forage_consumption [, grep("sucrose_mass_out", colnames(forage_consumption))]
  
  sucrose_mass_out_values <- cbind(colony_number = forage_consumption [, grep("colony_number",
                                                                              colnames(forage_consumption))], sucrose_mass_out_values)
  
  zero_or_not <- sucrose_mass_out_values == 0
  
  any_zero_in_row <- apply(zero_or_not, 1, any)
  
  spill_rows <- sucrose_mass_out_values [any_zero_in_row,]
  
  spill_rows_col_num <- spill_rows$colony_number
  
  for (i in 1:length(spill_rows_col_num)) {
    
    pollen_consumption_long <- pollen_consumption_long [spill_rows_col_num [i] != pollen_consumption_long$colony_number, ]
    
  }

  # add the two insecticide columns from each forage type together
  insecticide_consumption <- cbind(data.frame(pollen_consumption_long$colony_number),
                                   data.frame(pollen_consumption_long$block),
                                   data.frame(pollen_consumption_long$treatment),
                                   data.frame(pollen_consumption_long$triad),
                                   data.frame(pollen_consumption_long$day),
                                   data.frame(pollen_consumption_long$number_of_workers_at_exposure_start),
                                   data.frame(pollen_consumption_long$insecticide_consumption),
                                   data.frame(sucrose_consumption_long$insecticide_consumption),
                                   data.frame(sucrose_consumption_long$insecticide_consumption + pollen_consumption_long$insecticide_consumption))
  
  colnames(insecticide_consumption) <- c("colony_number", "block", "treatment", "triad", "day", "number_of_workers_at_exposure_start",
                                         "insecticide_consumption_pollen", "insecticide_consumption_sucrose", "insecticide_consumption_combined")
  
  # subset by colony, sum day insecticide consumption values to return total insecticide consumption
  total_dose_df <- tibble()
  
  for (i in 1:length(unique(insecticide_consumption$colony_number))) {
    
    by_colony <- insecticide_consumption [insecticide_consumption$colony_number == unique(insecticide_consumption$colony_number) [i], ]
    
    total_dose <- cbind(data.frame(unique(insecticide_consumption$colony_number) [i]), data.frame(sum(by_colony$insecticide_consumption_combined)))
    
    total_dose_df <- rbind(total_dose_df, total_dose)
    
  }
  
  colnames(total_dose_df) <- c("colony_number", "total_insecticide_dose")
  
  # re-associate total_insecticide_dose with explanatory variables
  total_insecticide_dose <- col_name_df_combiner_function(dataframe_1 = total_dose_df, dataframe_2 = sucrose_consumption_df)
  
  total_insecticide_dose <- total_insecticide_dose [, -grep("sucrose|summed", colnames(total_insecticide_dose))]
  
  # remove controls colonies for which the output is nonsense
  total_insecticide_dose <- total_insecticide_dose [total_insecticide_dose$treatment != "control",]
  
  return(total_insecticide_dose)
    
}

total_insecticide_dose_df <- total_insecticide_dose_function(forage_consumption = forage_consumption, 
                                                          pollen_consumption_long = pollen_consumption_long,
                                                          sucrose_consumption_long = sucrose_consumption_long,
                                                          sucrose_consumption_df = sucrose_consumption_df)
```

Create male_mass_df

```{r}

```

Write the final csvs to the output file

```{r}
# queen_survival
write_csv(queen_survival_days_df, "output/queen_survival.csv")

# colony_survival
write_csv(colony_survival_days_df, "output/colony_survival.csv")

# worker_number
write_csv(worker_number_df, "output/worker_number.csv")

# nectar_pot_number
write_csv(nectar_pot_number_df, "output/nectar_pot_number.csv")

# male_number
write_csv(total_reproductive_output_df, "output/total_male_number.csv")

# sucrose_consumption
write_csv(sucrose_consumption_df, "output/sucrose_consumption.csv")

# pollen_consumption
write_csv(pollen_consumption_df, "output/pollen_consumption.csv")

# insecticide_consumption
write_csv(total_insecticide_dose_df, "output/total_insecticide_dose.csv")

# male_mass
```
