---
title: "Colony Survival"
author: "Guy Mercer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Two sources I used for below info 

[link 1](https://rpubs.com/daspringate/survival)
[link 2](https://thomaselove.github.io/432-notes/cox-regression-models-for-survival-data-example-1.html)

Hazard Function
Survival models can be viewed as consisting of two parts: the underlying hazard function, describing how the risk of event per time unit changes over time at baseline levels of covariates; and the effect parameters, describing how the hazard varies in response to explanatory covariates.

h(t|x) = h0(t) exp[B1.x]

h0(t) = is the baseline hazard, which is a non-parametric and unspecified value which depends on t but not on x.

exp[B1.x] = how the baseline hazard is modified as x changes. 

It is a semi-parametric model:

The baseline hazard function is unspecified
The effects of the covariates are multiplicative (proportional assumption)
Doesn't make arbitrary assumptions about the shape/form of the baseline hazard function

Proportionality Assumption
In a proportional hazards model, the unique effect of a unit increase in a covariate is multiplicative with respect to the hazard rate. For example, taking a drug may halve one’s hazard rate for a stroke occurring, or, changing the material from which a manufactured component is constructed may double its hazard rate for failure.

Covariates multiply the hazard by some constant
e.g. a drug may halve a subjects risk of death at any time
*The effect is the same at any time point*

Accounting for non-proportional hazards
Include an interaction with time for the variables:
This factors time out of the main effect
Only use if it makes sense to have a linear interaction between the covariate and time (look at the graphs!)

Import data

```{r}
col_surv <- read.csv("./input/colony_survival.csv")

# ensure variables are in correct class
col_surv$block <- as.factor(col_surv$block)
col_surv$treatment <- as.factor(col_surv$treatment)
col_surv$triad <- as.factor(col_surv$triad)
col_surv$number_of_workers_at_exposure_start <- as.numeric(col_surv$number_of_workers_at_exposure_start)
col_surv$wax_moth <- as.factor(col_surv$wax_moth)
col_surv$colony_survival_days <- as.numeric(col_surv$colony_survival_days)
```

I used the example dataset "ovarian" in the survival package to set up my data. In a really old (1999) version of the survival package vignette I found the definition of fustat and futime. 

futime: The number of days from enrollment until death or censoring, whichever
came first.

fustat: An indicator of death (1) or censoring (0).

Therefore my data should have columns for survivalday and death

In death column death = 1, censoring = 0. As they all died eventually simply a column of 1s.

```{r}
col_surv$col_death <- 1
```

Remove colony 57 as the queen was accidentally killed.

```{r}
col_surv <- col_surv [col_surv$colony_number != "57",]
```

Begin the analysis. Attempt to start with:

col_surv ~ treatment + worker_num + wax_moth + worker_num:treatment + wax_moth:treatment + (1 | batch/triad)

```{r}
library(survival)
library(coxme)

cfit1 <- coxme(Surv(colony_survival_days, col_death) ~ treatment + number_of_workers_at_exposure_start + wax_moth + 
                 number_of_workers_at_exposure_start:treatment + wax_moth:treatment + (1 | block/triad),
               data=col_surv)

# summary(cfit1)

# model selection
cfit2 <-  coxme(Surv(colony_survival_days, col_death) ~ treatment + number_of_workers_at_exposure_start + wax_moth + 
                 number_of_workers_at_exposure_start:treatment + (1 | block/triad),
               data=col_surv)

anova(cfit1, cfit2)

cfit3 <- coxme(Surv(colony_survival_days, col_death) ~ treatment + number_of_workers_at_exposure_start + wax_moth + 
                 wax_moth:treatment + (1 | block/triad),
               data=col_surv)

anova(cfit1, cfit3)

# removed wax_moth:treatment
cfit4 <-  coxme(Surv(colony_survival_days, col_death) ~ treatment + number_of_workers_at_exposure_start + wax_moth + (1 | block/triad),
               data=col_surv)

anova(cfit2, cfit4)

cfit5 <-  coxme(Surv(colony_survival_days, col_death) ~ treatment + number_of_workers_at_exposure_start + 
                 number_of_workers_at_exposure_start:treatment + (1 | block/triad),
               data=col_surv)

anova(cfit2, cfit5)

# removed wax_moth
cfit6 <-  coxme(Surv(colony_survival_days, col_death) ~ treatment + number_of_workers_at_exposure_start + (1 | block/triad),
               data=col_surv)

anova(cfit5, cfit6)

# removed number_of_workers_at_exposure_start:treatment
cfit7 <-  coxme(Surv(colony_survival_days, col_death) ~ treatment + (1 | block/triad),
               data=col_surv)

anova(cfit6, cfit7)

# dropped number_of_workers_at_exposure_start
cfit8 <-  coxme(Surv(colony_survival_days, col_death) ~ (1 | block/triad),
               data=col_surv)

anova(cfit7, cfit8)

# treatment is not significant either but leave in model as this is the variable of interest. 
```

Summary of treatment only model 

```{r}
cfit_treat <- cfit7

summary(cfit_treat)

cfit_treat_no_ran <- coxph(Surv(colony_survival_days, col_death) ~ treatment,
               data=col_surv)

summary(cfit_treat_no_ran)
```

The coefficients in a Cox regression relate to hazard; a positive coefficient indicates a worse prognosis and a negative coefficient indicates a protective effect of the variable with which it is associated.

Model Validation: 

zph() checks for proportionality assumption, by using the Schoenfeld residuals against the transformed time. Having very small p values indicates that there are time dependent coefficients which you need to take care of. That is to say, the proportionality assumption does not check linearity - the Cox PH model is semi parametric and thus makes no assumption as to the form of the hazard. The proportionality assumption is that the hazard rate of an individual is relatively constant in time, and this is what cox.zph() tests.

A good [resource on checking assumptions of proportional hazards models](http://www.sthda.com/english/wiki/cox-model-assumptions#testing-proportional-hazards-assumption)

```{r}
# proportional hazards
zp1 <- cox.zph(cfit_treat)
zp1

# plot(zp1[1], resid=TRUE)

library(survminer)
ggcoxzph(zp1)
```

Testing for influential observations - visualizing the deviance residuals. The deviance residual is a normalized transform of the martingale residual. These residuals should be roughtly symmetrically distributed about zero with a standard deviation of 1.

1. Positive values correspond to individuals that “died too soon” compared to expected survival times.
2. Negative values correspond to individual that “lived too long”.
3. Very large or small values are outliers, which are poorly predicted by the model.

Unfortunately coxme objects can't be used for some of these check. Use cfit_treat_no_ran instead. Coefficients are pretty similar to random effects model. 

```{r}
ggcoxdiagnostics(cfit_treat_no_ran, type = "deviance",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```

The dfbeta plot shows the effect dropping an observation has on the coefficients. Shows there are two points that affect the flup coefficient strongly. Removing these two points only increases flup's "protective" effect. Also, I have no reason to exclude them from an experimental standpoint. For these reasons, leave them in. 

```{r}
ggcoxdiagnostics(cfit_treat_no_ran, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```

non-linearity is not an issue with categorical variables. In future refer to [guide](http://www.sthda.com/english/wiki/cox-model-assumptions#testing-influential-observations) if this check is required. 

```{r}
library(survminer)

# log rank test between KM survival curves.
# surv <- Surv(time = col_surv$colony_survival_days, event = col_surv$col_death)
# survdiff(surv ~ col_surv$treatment)

# risk.table = TRUE,  risk.table.height = 0.25

fit1 <- survfit(Surv(colony_survival_days, col_death) ~ treatment, data=col_surv)


# for prettier graphs
p <- ggsurvplot(fit1,
                data = col_surv,
                conf.int = TRUE,
                palette = c("#117733", "#332288", "#AA4499"),
                legend.labs = c("Control", "FPF", "Sivanto"),
                xlab = "Time (days)",
                ylab = "Probability of Colony Survival",
                legend.title = "",
                title = "Colony Survival")

p <- p + theme_survminer(font.main = c(24, "bold", "black"),
                         font.x = c(18, "bold", "black"),
                         font.y = c(18, "bold", "black"),
                         font.tickslab = c(16, "plain", "black"),
                         font.legend = c(16, "plain", "black"))

p
```

[Colour Palette] (https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible)

```{r}
final_model <- coxme(Surv(colony_survival_days, col_death) ~ treatment + (1 | block) + (1 | campus_location),
               data=col_surv)

summary(final_model)
```

